### 2018-03-14 paper/user-model.tex

To this end, we use a Hidden Markov Model (HMM). 
The \emph{hidden} Markovian process dictates the user's progression among the three proposed locations.
The \emph{observable} process emits ``toggle'' actions on the user's devices, based on probabilities associated with each location.
For instance, the probability that the user switches her home computer on or off (i.e. \emph{toggles}) is null while she is at work or outside.

Such a statistical model is overly simplistic (e.g. it does not encapsulate the time or the weekday), but it is on purpose: simplicity is bliss.
Given the lack of real-world data for our testbed, building a more complex and realistic user model would add no value to our proposal: 
our goal is to see devices exhibiting frequent disconnections and reconnections (or \emph{churn}).
Our simple model being able to put extravagant stress on our system, it is sufficient to access \name's resilience.

>>>>

To this end, we use a Hidden Markov Model (HMM). 
The \emph{hidden} Markovian process $S=S_1,\dots,S_T$ dictates the user's progression among the three proposed locations.
The \emph{observable} process $O=O_1,\dots,O_T$ emits ``toggle'' actions on the user's devices, based on probabilities associated with each location.
For instance, the probability that the user switches her home computer on or off (i.e. \emph{toggles}) is null while she is at work or outside.

Such a statistical model is overly simplistic (e.g. it does not encapsulate the time or the weekday), but it is on purpose: simplicity is bliss.
Given the lack of real-world data for our testbed, building a more complex and realistic user model would add no value to our proposal: 
our goal is to see devices exhibiting frequent disconnections and reconnections (also called \emph{churn}).
Our simple model being able to put extravagant stress on our system, it is sufficient to access \name's resilience.

\subsubsection{Formal definition} % (fold)
\label{ssub:formal_definition}

\begin{figure}[t]
\centering
\vspace{-1em}

$$A =
\kbordermatrix{
      & W            & O            & H            \cr
    W & \sfrac{2}{3} & \sfrac{1}{3} & 0            \cr
    O & \sfrac{1}{3} & \sfrac{1}{3} & \sfrac{1}{3} \cr
    H & 0            & \sfrac{1}{3} & \sfrac{2}{3} \\[0.3em]
}, \;
B = 
\kbordermatrix{
      & W    & O    & H    \cr
    n & 0.1  & 0.45 & 0.1  \cr
    p & 0.45 & 0.55 & 0.45 \cr
    w & 0.45 & 0    & 0    \cr
    h & 0    & 0    & 0.45 \\[0.3em]
}$$

\includegraphics[width=0.85\columnwidth]{figures/hmm.pdf}
\caption{ \label{fig:hmm} Example Hidden Markov Model (HMM) of a user's behavior.}
\end{figure}

We consider a single user, owning a set of $\mathcal{D}$ devices.
Time is discretized here, and the user's activity is sequential: she can do zero or one action per ``round''of $T$s.
A HMM $\lambda_N$, comprising a set of states $\mathcal{S}$ and a set of observable events $\mathcal{O}$, is defined as follows:

$$\text{HMM}:\;\lambda_N=(\Pi, A, B)$$

where 
\begin{itemize}
	\item $N$ is the number of states in the system:
	$N = \left| \mathcal{S} \right|$;

	\item $\Pi=\left\{ \pi_i\right\}_{i\in\mathcal{S}}$ is the initial state distribution:\\
	$\pi_i=P[S_1=i]$;

	\item $A = \left\{ a_{ij}\right\}_{(i,j)\in\mathcal{S}^2}$ is the transition probability matrix:\\
	$a_{ij}=P[S_t=j \mid S_{t-1}=i]$;

	\item $B = \left\{ b_{ik}\right\}_{(i,k)\in\mathcal{S}\times\mathcal{O}}$ is the matrix of states conditional densities:
	$b_{ik} = P[O_t=k \mid S_t = i]$.
\end{itemize}

In our case, the states are the different locations where the user is susceptible to go.
We considered the three following locations: $\mathcal{S}=\left\{ \mathit{work}, \mathit{outside}, \mathit{home} \right\}$. 
There is one action per device $d$ owned by the user, plus a $\mathit{none}$ action when the user does not do anything: 
$\mathcal{O} = \left\{ \mathit{toggle}_d \right\}_{d\in \mathcal{D}}$.
$\mathit{toggle}_d$ means that the user turned her device $d$ on or off, depending on the device's previous state.

In figure~\ref{fig:hmm}, we show an example HMM of a user's behavioral HMM, along with its graphical representation (the initial state distribution $\Pi$ was left out for simplicity). 
The represented user possesses a phone, that she carries around with her, a home computer, that is only accessible from her home, and a workstation, located at her workplace.

\begin{figure}[t]
\centering
\includegraphics[width=\columnwidth]{figures/sample_usage.pdf}

\caption{\label{fig:sample_usage}Sample device usage using the previous HMM specifications. 
%The bottom line reads the HMM's states: \textbf{W}ork, \textbf{O}utside, and \textbf{H}ome, while the dark lines read when the devices (\textbf{h}ome computer, \textbf{p}hone, and \textbf{w}orkstation) are online.
}
\end{figure}

\commentAL{Stupid of me to use same abbreviations for toggle and usage events...}

Figure~\ref{fig:sample_usage} plots a sample device usage timeline using the previous model. 
Each device starts offline; the user can toggle one device per round (or do nothing).
In this figure, we show the device \emph{usage}, that is a consequence of the $\mathit{toggle}$ events defined by the HMM.
As a consequence, several devices can be online at the same time, as happens in the figure.

% subsubsection formal_definition (end)
